\LARGE \bf When Children Teach a Robot to Write: An Autonomous Teachable Humanoid Which Uses Simulated Handwriting

### 2014-12-12 main.tex
final changes throughout

Within this article, Section \ref{sec:learningAlgorithm} presents the novel work in the area of
artificial intelligence to develop a learning algorithm suitable for a teachable
agent in the context of handwriting. Section \ref{sec:robotWriting} details the
extension of this algorithm to an embodied robotic learning agent, including the
new approach for achieving simulated fine motor skills on commercially
affordable humanoid robots such as the {\sc nao}. Section \ref{sec:experiment} explores the 
contributions made to the study of human-robot interaction, in discussing the
use of the system with primary school children and its potential as a tool for
addressing wider pedagogical research questions in education. Finally, Section \ref{sec:futureWork} addresses the challenges which are faced in extending this system to a level suitable for long-term studies, and Section \ref{sec:conclusions} concludes by reiterating the content and impact of the article's contributions.

@@

Teachable computer-based agents have previously been used to encourage the
``protégé effect'', wherein students invest more effort into learning when it is
for a teachable agent than for themselves~\cite{Chase2009}.  As we are concerned
with learning of a physical skill, the learning agent developed is embodied in a
humanoid robot which is capable of physically demonstrating handwriting
trajectories to its child learning partner.  This is supported by the potential
for motor mimicry to yield significant improvements in handwriting interventions
in which letter formations are demonstrated to
participants~\cite{Berninger1997}. Furthermore, when compared to screen-based
agents, robotic partners have been shown in some contexts to increase users'
compliance with tasks~\cite{Bainbridge2011}, maintain more effective long-term
relationships~\cite{Kidd2008}, and produce greater learning gains when acting as
tutors~\cite{Leyzberg2012} that can be adapted to each
learner~\cite{Leyzberg2014}.

Robots have been used as teachers or social partners to promote children's
learning in a range of contexts, most commonly related to language
skills~\cite{Han2010}, and less often to physical skills (such as
calligraphy~\cite{Matsui2013}).  Looking at the converse (humans \emph{teaching}
robots), Werfel notes in~\cite{Werfel2014} that most of the work focuses on the
robot's benefits (in terms of language~\cite{Saunders2010} or
physical~\cite{Mulling2013} skills, for example) rather than the learning
experienced by the human tutor itself.  Our work concentrates on this latter
aspect: by demonstrating handwriting to a robot, we aim at improving the
\emph{child} performances. Note that our work must be distinguished from
traditional ``learning from demonstration'' approaches, as the agent we present
is only simulating fine motor skills for interaction purposes.

@@

A parameterisation of letters and their deformities was sought such 
that different quality shapes can be generated, depending on the parameters input to
the letter models. This allows us to configure the system such as to improve its writing 
by modifying the parameters based on feedback from the 
reinforcement learning partner (Section~\ref{sec:learningAlg}).

@@

We use statistical shape modelling as an approach for a shape model which can
appropriately represent realistic variations in shapes. Statistical shape
modelling is an application of principle component analysis (PCA), where a
linear transform which de-correlates data vectors is found~\cite{Stegmann2002}
and allows for dimensionality reduction.

PCA is performed on a set of letter paths captured from a digital pen,
using the UJI Pen Characters 2 dataset \cite{Llorens2008} with 120 instances of
each letter (2 repetitions from 60 adult users). While it may be appropriate in future work to
identify the location of salient features of the shapes which are robust to 
unanticipated user input (such as shapes drawn backwards),
the features are currently taken as $n$ uniformly spaced points along the shape
path. The
points are arranged into an observation vector presented in (\ref{eq:obsVec}),
where $x_i$ and $y_i$ represent the coordinates of each of the points along the
path. The observation shapes are normalised to have a unit maximum dimension 
and {\bf 0} mean.

@@

Equation (\ref{eq:projection}) represents the projection from the original 
$2n$-dimensional feature space $(n=70)$ to a reduced $N$-dimensional space, where ${\bf p}$
contains the coordinates in the $N$-dimensional space with ${\bf 0}$-origin,
calculated as in (\ref{eq:paramCalc}). 
${\bf \Phi }$ is
an orthogonal $2n\times N$ matrix composed of the eigenvectors ${\bf v}_i$ corresponding to
the largest $N$ eigenvalues ($\lambda_i$) of the covariance matrix of the
observations, as shown in (\ref{eq:transformCalc}) \cite{Stegmann2002}. 
If there is correlation between the points in the observations, there will be
eigenvalues of the covariance matrix which are close to zero. As such, removing
the associated eigenvectors from ${\bf \Phi}$ allows for dimensionality
reduction with minimal impact.

@@

PCA is performed on all of the paths of a particular allograph in the
dataset individually, to reduce the $2n$-dimensional space for that shape to one 
with $N=10$ dimensions.
As such,
each shape may be approximated by the mean shape of the allograph plus a sum of the top 10
eigenvectors, weighted by the parameter vector ${\bf p}$. 
%Indeed, the amount of variance in the dataset
%that is explained by each eigenvector ${\bf v}_i$ is $\sqrt{\lambda_i}$, and
%therefore the proportion of the variance explained by each dimension/eigenvector
%is as shown in (\ref{eq:proportionVar}).
%
%\begin{equation}\label{eq:proportionVar}
%\%var_i = \frac{\sqrt{\lambda_i}}{\sum_{i=1}^{N}\sqrt{\lambda_i}}\times 100\%
%\end{equation}

(\ref{eq:projection}) may also be used
to generate new shapes based on the parameters ${\bf p}$ which are used. ${\bf
p}={\bf 0}$ will yield the mean shape, and variations to each of the $N$ values
in ${\bf p}$ will cause a change in the shape represented by the corresponding
eigenvector (Figure \ref{fig:deviations_sPrint}). 
For the dataset presented in Figure \ref{fig:deviations_sPrint}, the
eigenvectors associated with the 3 largest eigenvalues explain 78.5\% of the
variance in the dataset, illustrating the capability of the statistical shape
modelling approach to produce compact parameterisation of shapes.

%Clusters within paths of a particular letter have been identified using K-means
%clustering to group different styles of writing the letter (\emph{allographs}),
%such that the parameters in the resulting model are only representative of the
%specific allograph of the letter in the dataset. 
% * <deanna.m.hood@gmail.com> 2014-09-03T13:05:23.576Z:
%
% in the end this was pointless so maybe not worth mentioning
%

@@

Interestingly, although the parameters are the result of an \emph{unsupervised} 
shape analysis, they still represent variations which could have
been intuitively identified by a manual parameterisation. For example, for the
model shown in Figure \ref{fig:deviations_sPrint}, the parameters may represent
the height of the top half of the letter compared to the bottom half, the width
of the overall shape, etc. The ability to generate varied levels of deformations
which may be ascribed descriptive interpretations (not just numerical) is an
advantage of this method, given its intended use with humans. It is, for instance,
possible for a teacher to create initial letters with a particular feature (a wide
`s' or a `d' with a large loop, for example).

@@

As explained, new letters can be generated by varying the parameter values 
for a shape model in accordance with (\ref{eq:projection}). By choosing
parameter values which lie within the observed range in the dataset, it is
possible to produce letters which are more likely to be reasonable looking.
When the parameter values are outside of the range observed in the dataset, they
are less likely to represent shapes from the dataset of adult-written letters, 
and as a result are more likely to represent poor shapes.
Figure \ref{fig:sampleLetters} illustrates sample letters generated from the
models of `e' and `g' by selecting random values for the first 5 parameters 
from a distribution with
standard deviation of $3\sqrt\lambda_i$ rather than the $\sqrt\lambda_i$
standard deviation observed in the dataset.

@@

For the statistical shape model employed by the system, the method for
responding to user demonstrations is to move the learning algorithm's parameters
towards those of the demonstration. In the results presented in this work, the
linear update equation shown in (\ref{eq:update}) is used, where ${\bf p}$ is the
learned parameter vector at time step $k$, and $\alpha$ is the learning rate,
between 0 and 1.

@@

It is possible that parameters ${\bf p}^{(k)}$ and ${\bf p}_{demo}$, which
individually yield acceptable shapes, produce parameters ${\bf p}^{(k+1)}$
which yield an unacceptable shape. This is especially true if the demonstration
shape is of a different style to that learnt at time $k$, as there are no
restrictions imposed on parameter values. However, the proposed method for adapting to
the demonstration shapes has the advantage of being able to recover from such a situation: with further
demonstrations of the same letter, the system would eventually approach the
demonstration shape (see Figure \ref{fig:eDemo}. As a result, the event of poor-looking 
intermediate letters
would not limit the interaction later proposed in Section \ref{sec:experiment}
in a technical sense, but it may influence the user's \emph{perception} of the learning
agent. It remains to be seen if it is necessary to avoid such an event
mathematically.

@@

The {\sc nao} V4 humanoid robot, which has been developed by Aldebaran Robotics and 
purposely designed to look approachable \cite{Gouaillier2008}, is used for
this work. It is a commercially affordable biped robot, 58cm tall, with 25
degrees of freedom, two cameras, speech capabilities and the ability to
autonomously execute a range of tasks.

@@

To enable the robot's `writing' to display while the robot is tracing
trajectories, ROS is used for the communication between the devices,
including the Android tablet\footnote{For more information about ROS on Android
devices see \url{http://wiki.ros.org/android}}. As a result, aspects of the
networking between the tablet and the robot, such as the overheads associated
with connections, ports, message formats, etc. have been simplified.

@@

The user demonstrations are received by the
interaction controller running on a desktop computer. It is responsible for
getting the {\sc nao} to prompt and respond appropriately to feedback received using
a finite state machine to manage the interaction stage and various system
inputs. In the context of learning handwriting, an additional input to the system 
is a word from which the letters are to be learned, which is detected by a fiducial 
marker on the card displaying the word.
The controller provides inputs to the learning algorithm including the word to learn
and the user demonstrations, by inferring the letter which the demonstrations are 
intended for based on their position on the tablet. The output shapes from the 
learning algorithm are sent by the controller to the devices which write them.

@@

    \item The participant shows the robot one of seven different 3-letter words
        to write, are made up of 7 possible letters (`c', `e',
        `n', `o', `s', `u', `w'). Fiducial markers which are printed on the word
        cards allow them to be detected with the robot's camera.

@@

\subsection{Outcomes of Preliminary Study 2}
The study at the second school involved 21 french-speaking students aged 7-8 years. 7 children
interacted with the robot individually and 7 sessions included the remaining
14 students interacting with the robot in pairs.
Initial parameters were drawn from a range purposely chosen to generate shapes 
for letters 'e' and 's' which would elicit correction. For the other letters, 
parameters were fixed, generating shapes which some groups still chose to correct 
(e.g. 13/14 for 'n' and 2/14 for 'c'). The duration of the sessions
was between 8 and 15 minutes, with an average of 11.4 minutes (SD = 2.3).

@@

A conclusion drawn in a systematic review of handwriting intervention studies~\cite{Hoy2011} 
is that any of the studies considered which involved 
fewer than two practice sessions per week and fewer than a total of 20 practice 
sessions, including homework, were not found to demonstrate effective results. This 
highlights the necessity to engage students in an interaction which will be sustainable
 over the long-term if 
we want to address research questions which involve the measurement of learning gains. 
Several challenges are raised in developing such long-term capabilities for the system,
 from both technical and sociological points of view. %not sure this makes sense

@@

With some of these challenges addressed, further steps can be made towards 
answering the overarching question of what impact the 
addition of such a teachable robotic agent would have on handwriting interventions, including the effect on the 
participants' self-esteem, motivation, and learning gains. % does it impact the
%motivation of the participants, their self-esteem, and/or the learning gains
%that they experience? We are currently considering running a long-term study
%that would provide better evidences to answer these questions.

@@

\addtolength{\textheight}{-.5cm}   % This command serves to balance the column lengths
                                  % on the last page of the document manually. It shortens
                                  % the textheight of the last page by a suitable amount.
                                  % This command does not take effect until the next page
                                  % so it should come on the page before the last. Make
                                  % sure that you do not shorten the textheight too much.

@@

We believe that this article introduces three noteworthy contributions: an 
innovative application of artificial intelligence for the learning of 
hand-written letters suitable for educative purposes; a method for a 
commercially affordable humanoid, in particular the {\sc nao} robot, to exhibit
simulated fine motor capabilities; and a working autonomous interaction 
suitable for engaging children in complex human-robot interactions, such as 
teacher-learner social interactions and learning from demonstration.

>>>>

Within this article, Section \ref{sec:learningAlgorithm} presents the novel work in the area of
artificial intelligence to develop a learning algorithm suitable for a teachable
agent in the context of handwriting. Section \ref{sec:robotWriting} details the
extension of this algorithm to an embodied robotic learning agent, including the
new approach for achieving simulated fine motor skills on commercially
affordable humanoid robots such as the {\sc nao}. Section \ref{sec:experiment} explores the 
contributions made to the study of human-robot interaction, in discussing the
use of the system with primary school children and its potential as a tool for
addressing wider pedagogical research questions in education. Finally, Section \ref{sec:futureWork} addresses the challenges which are faced in extending this system to a level suitable for long-term studies, and Section \ref{sec:conclusions} concludes by reiterating the impact of the article's contributions.

@@

Teachable computer-based agents have previously been used to encourage the
``protégé effect'', wherein students invest more effort into learning when it is
for a teachable agent than for themselves~\cite{Chase2009}.  As we are concerned
with learning of a physical skill, the learning agent developed is embodied in a
humanoid robot which is capable of physically demonstrating handwriting
trajectories to its child learning partner.  This is supported by the potential
for motor mimicry to yield significant improvements in handwriting interventions
in which letter formations are demonstrated to
participants~\cite{Berninger1997}. Furthermore, when compared to screen-based
agents, robotic partners have been shown in some contexts to increase users'
compliance with tasks~\cite{Bainbridge2011}, maintain more effective long-term
relationships~\cite{Kidd2008}, and produce greater learning gains when acting as
tutors~\cite{Leyzberg2012}.

Robots have been used as teachers or social partners to promote children's
learning in a range of contexts, most commonly related to language
skills~\cite{Han2010}, and less often to physical skills (such as
calligraphy~\cite{Matsui2013}).  Looking at the converse (humans \emph{teaching}
robots), Werfel notes in~\cite{Werfel2014} that most of the work focuses on the
robot's benefits (in terms of language~\cite{Saunders2010} or
physical~\cite{Mulling2013} skills, for example) rather than the learning
experienced by the human tutor themselves.  Our work concentrates on this latter
aspect: by demonstrating handwriting to a robot, we aim at improving the
\emph{child's} performance. Note that our work must be distinguished from
``learning from demonstration'' approaches to robots learning physical skills, 
as the agent we present is only simulating fine motor skills for interaction 
purposes.

@@

A parameterisation of letters and their deformities is used such 
that different quality shapes can be generated, depending on the parameters input to
the letter models. This allows us to configure the system to improve its writing 
by modifying the parameters based on feedback from the 
reinforcement learning partner (Section~\ref{sec:learningAlg}).

@@

We use statistical shape modelling for generating a shape model which can
appropriately represent realistic variations in shapes. Statistical shape
modelling is an application of principle component analysis (PCA), where a
linear transform which de-correlates data vectors is found~\cite{Stegmann2002}
and allows for dimensionality reduction.

PCA is performed on a set of letter paths captured from a digital pen,
using the UJI Pen Characters 2 dataset \cite{Llorens2008} with 120 instances of
each letter (2 repetitions from 60 adult users). While it may be appropriate in future work to
identify the location of salient features of the shapes which are robust to 
unanticipated user input (such as shapes drawn backwards),
the features are currently taken as $n=70$ uniformly spaced points along the shape
path. The
points are arranged into an observation vector presented in (\ref{eq:obsVec}),
where $x_i$ and $y_i$ represent the coordinates of each of the points along the
path. The observation shapes are normalised to have a unit maximum dimension 
and {\bf 0} mean.

@@

Equation (\ref{eq:projection}) represents the projection from the original 
$2n$-dimensional feature space to a reduced $N$-dimensional space, where ${\bf p}$
contains the coordinates in the $N$-dimensional space with ${\bf 0}$-origin,
calculated as in (\ref{eq:paramCalc}). 
${\bf \Phi }$ is
an orthogonal $2n\times N$ matrix composed of the eigenvectors ${\bf v}_i$ corresponding to
the largest $N$ eigenvalues ($\lambda_i$) of the covariance matrix of the
observations~\cite{Stegmann2002}, as shown in (\ref{eq:transformCalc}). 
If there is correlation between the points in the observations, there will be
eigenvalues of the covariance matrix which are close to zero. As such, removing
the associated eigenvectors from ${\bf \Phi}$ allows for dimensionality
reduction with minimal impact.

@@

PCA is performed on all of the paths of a particular allograph in the
dataset individually, to reduce the $2n$-dimensional space for that shape to one 
with $N=10$ dimensions.
Each shape is then approximated by the mean shape of the allograph plus a sum of the top 10
eigenvectors, weighted by the parameter vector ${\bf p}$. 
%Indeed, the amount of variance in the dataset
%that is explained by each eigenvector ${\bf v}_i$ is $\sqrt{\lambda_i}$, and
%therefore the proportion of the variance explained by each dimension/eigenvector
%is as shown in (\ref{eq:proportionVar}).
%
%\begin{equation}\label{eq:proportionVar}
%\%var_i = \frac{\sqrt{\lambda_i}}{\sum_{i=1}^{N}\sqrt{\lambda_i}}\times 100\%
%\end{equation}

Equation (\ref{eq:projection}) may also be used
to generate new shapes based on the parameters ${\bf p}$ which are used. ${\bf
p}={\bf 0}$ will yield the mean shape, and variations to each of the $N$ values
in ${\bf p}$ will cause a change in the shape represented by the corresponding
eigenvector (Figure \ref{fig:deviations_sPrint}). 
For the dataset presented in Figure \ref{fig:deviations_sPrint}, the
eigenvectors associated with the 3 largest eigenvalues explain 78.5\% of the
variance in the dataset, illustrating the capability of the statistical shape
modelling approach to produce compact parameterisation of shapes.

@@

Interestingly, although the parameters are the result of an \emph{unsupervised} 
shape analysis, they still represent variations which could have
been intuitively identified by a manual parameterisation. For example, for the
model shown in Figure \ref{fig:deviations_sPrint}, the parameters may represent
the height of the top half of the letter compared to the bottom half, the width
of the overall shape, etc. The ability to generate varied levels of deformations
which may be ascribed descriptive interpretations (not just numerical) is an
advantage of this method, given its intended use with humans. It is, for instance,
possible for a teacher to create letters -- which will be used as a starting point
for the system -- with a particular feature (a wide
`s' or a `d' with a large loop, for example).

@@

As explained, new letters can be generated by varying the parameter values 
for a shape model in accordance with (\ref{eq:projection}). By choosing
parameter values which lie within the observed range in the dataset, it is
possible to produce letters which are more likely to be reasonable looking.
When the parameter values are outside of the range observed in the dataset, they
are less likely to represent shapes from the dataset of adult-written letters, 
and as a result are more likely to represent poor shapes.
Figure \ref{fig:sampleLetters} illustrates sample letters generated from the
models of `e' and `g' by selecting random values for the first 5 parameters 
from a distribution with
standard deviation of $3\sqrt\lambda_i$, rather than the $\sqrt\lambda_i$
standard deviation observed in the dataset.

@@

The method we employ for
responding to user demonstrations is to move the learning algorithm's parameters
towards those of the demonstration. In the results presented in this work, the
linear update equation shown in (\ref{eq:update}) is used, where ${\bf p}$ is the
learned parameter vector at time step $k$, and $\alpha$ is the learning rate,
between 0 and 1.

@@

It is possible that parameters ${\bf p}^{(k)}$ and ${\bf p}_{demo}$, which
individually yield acceptable shapes, produce parameters ${\bf p}^{(k+1)}$
which yield an unacceptable shape. This is especially true if the demonstration
shape is of a different style to that learnt at time $k$ (see 
Figure~\ref{fig:eDemo}), as there are no
restrictions imposed on parameter values. However, the proposed method for adapting to
the demonstration shapes has the advantage of being able to recover from such a situation: with further
demonstrations of the same letter, the system would eventually approach the
demonstration shape. As a result, the event of poor-looking 
intermediate letters
would not limit the interaction later proposed in Section \ref{sec:experiment}
in a technical sense, but it may influence the user's \emph{perception} of the learning
agent. It remains to be seen if it is necessary to avoid such an event
mathematically.

@@

The {\sc nao} V4 humanoid robot, which has been purposely designed by 
Aldebaran Robotics to look approachable \cite{Gouaillier2008}, is used for
this work. It is a commercially affordable biped robot, 58cm tall, with 25
degrees of freedom, two cameras, speech capabilities and the ability to
autonomously execute a range of tasks.

@@

To enable the robot's `writing' to display while the robot is tracing
trajectories, ROS is used for the communication between the devices,
including the Android tablet\footnote{For more information about ROS on Android
devices see \url{http://wiki.ros.org/android}}. As a result, aspects of the
networking between the tablet and the robot, such as the overheads associated
with connections, ports, etc. have been simplified.

@@

The user demonstrations are received by the
interaction controller running on a desktop computer. It is responsible for
getting the {\sc nao} to prompt and respond appropriately to feedback received using
a finite state machine to manage the interaction stage and various system
inputs. In the context of learning handwriting, an additional input to the system 
is a word from which the letters are to be learned, which is detected by a fiducial 
marker on the card displaying the word.
The controller provides inputs to the learning algorithm including the word to learn
and the user demonstrations, by inferring the letter which the demonstrations are 
intended for based on their position on the tablet. The output shapes from the 
learning algorithm are then sent again to the devices which write them.

@@

    \item The participant shows the robot one of seven different 3-letter words
        to write, made up of 7 possible letters (`c', `e',
        `n', `o', `s', `u', `w'). Fiducial markers which are printed on the word
        cards allow them to be detected with the robot's camera.

@@

\subsection{Outcomes of Preliminary Study 2}
The study at the second school involved 21 french-speaking students aged 7-8 years. 7 children
interacted with the robot individually and 7 sessions included the remaining
14 students interacting with the robot in pairs.
Initial parameters were drawn from a range purposely chosen to generate shapes 
for letters `e' and `s' which would elicit correction. For the other letters, 
parameters were fixed, generating shapes which some groups still chose to correct 
(e.g. 13/14 for `n' and 2/14 for `c'). The duration of the sessions
was between 8 and 15 minutes, with an average of 11.4 minutes (SD = 2.3).

@@

A conclusion drawn in a systematic review of handwriting intervention studies~\cite{Hoy2011} 
is that any of the studies considered which involved 
fewer than two practice sessions per week and fewer than a total of 20 practice 
sessions, including homework, were not found to demonstrate effective results. This 
highlights the necessity to engage students in an interaction which will be sustainable
 over the long-term if 
we want to address research questions which involve the measurement of learning gains. 
Several challenges are raised in developing such long-term capabilities for the system.

@@

Addressing these challenges will take us further towards answering if 
the addition of a teachable robotic agent to handwriting interventions 
would benefit the participants' self-esteem, motivation, and learning gains.

@@

\addtolength{\textheight}{-11.2cm}   % This command serves to balance the column lengths
                                  % on the last page of the document manually. It shortens
                                  % the textheight of the last page by a suitable amount.
                                  % This command does not take effect until the next page
                                  % so it should come on the page before the last. Make
                                  % sure that you do not shorten the textheight too much.

@@

We believe this article introduces three noteworthy contributions: an 
innovative application of artificial intelligence for the learning of 
hand-written letters suitable for educative purposes; a method for a 
commercially affordable humanoid, in particular the {\sc nao} robot, to exhibit
simulated fine motor capabilities; and a working autonomous interaction 
suitable for engaging children in complex human-robot interactions, such as 
teacher-learner social interactions and learning from demonstration.
